---
title: "covid_p_hacking"
author: "Maximiliana Bogan"
date: "9/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## P-hacking COVID-19

To show firsthand how p-hacking and overfitting are possible, we want you to show how these practices can lead to completely nonsensical results.

You can download a complete list of data on COVID-19 (coronavirus) by Our World in Data (https://ourworldindata.org/coronavirus). The data is updated daily and contains the lastest publicly available data on COVID-19 by country and by date. The data report the total cases, new cases, total deaths, new deaths, and hospitalization data of 233 countries and regions. Note that you are not expected to analyze the entire data. You may choose one or a few countries, or select one or some dates for analysis or for comparison.

The challenge is to build an analysis pipeline that produces a "significant" p-value for a relationship between COVID-19 cases and another variable, where the relationship is non-sensical, cannot possibly be causal, or could be argued either way. You may even simulate a fake variable as your key variable of interest. Prepare an Rmarkdown document with the results. At the end of the document write a paragraph to explain your "findings". As if you were in a debate team, pick on a subjective conclusion, and "cherry-pick" partial data to support your claim. Provide a non-statistical explanation for your group's fake result, and/or critique your statistcal approach and why your group got an apparently significant p-value.

#import libraries and read in file
```{r}
library(dplyr)
#read the Dataset sheet into “R”. The dataset will be called "data".
data <- read.csv("https://covid.ourworldindata.org/data/owid-covid-data.csv",
     na.strings = "",header=T)
```

#Finding correlations between average smokers and new COVID-19 cases in a country 
#male smokers and new cases
```{r}
results_m <- NULL
for(i in 1:length(unique(data$date))) {
  data_smoke_vax_m <- data[which((data$date==unique(data$date)[i]) & (data$continent=="Europe")),]
  data_smoke_vax_m <- data_smoke_vax_m %>%
    select(iso_code, 
           location, 
           date, 
           people_fully_vaccinated_per_hundred,
           new_cases,
           female_smokers, 
           male_smokers)
  if (sum(rowSums(!is.na(data_smoke_vax_m[,c(5,7)]))==2)>=5){
    results_m <- rbind(results_m, c(unique(data$date)[i],   #create result table w/ date...
    cor.test(data_smoke_vax_m[,5],data_smoke_vax_m[,7])$estimate, #... correlation
    cor.test(data_smoke_vax_m[,5],data_smoke_vax_m[,7])$p.value)) #... p-value
  }
}

results_m[which((as.numeric(results_m[,3])<=0.05)),]
```

We found a significant positive correlation between new COVID-19 cases and average number of male smokers in European countries on a number of dates throughout May, June, July, and August of 2020. Smokers are frequently directed to designated smoking areas, and may be more likely to congregate and spend extended periods in those locations without masks. This may have led to greater COVID-19 spread, and an increase in new COVID-19 cases in those countries. 

#female smokers and new cases - we ran both male and female smokers against new cases, but only found significant relationships with male smokers. 
```{r}
# results_f <- NULL
# for(i in 1:length(unique(data$date))) {
#   data_smoke_vax_f <- data[which((data$date==unique(data$date)[i]) & (data$continent=="Europe")),]
#   data_smoke_vax_f <- data_smoke_vax_f %>%
#     select(iso_code, 
#            location, 
#            date, 
#            people_fully_vaccinated_per_hundred,
#            new_cases,
#            female_smokers, 
#            male_smokers)
#   if (sum(rowSums(!is.na(data_smoke_vax_f[,5:6]))==2)>=5){
#     results_f <- rbind(results_f, c(unique(data$date)[i],   #create result table w/ date...
#     cor.test(data_smoke_vax_f[,5],data_smoke_vax_f[,6])$estimate, #... correlation
#     cor.test(data_smoke_vax_f[,5],data_smoke_vax_f[,6])$p.value)) #... p-value
#   }
# }
# 
# results_f[which((as.numeric(results_f[,3])<=0.05)),]
```